SI

### 2021-01-05 ms/manuscript.Rmd
Add SI plots: rel water content, DT controls for sporophytes

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22°C to 25°C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield [@WatkinsJr.2007c].

>>>>

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22°C to 25°C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].

### 2021-01-05 ms/si.Rmd
Add SI plots: rel water content, DT controls for sporophytes

---
title: "SI"
bibliography: 
  - references.bib
  - references_other.yaml
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research-si.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

@@

```{r load-data}
tar_load(
  names = c(
    "filmy_dt_chamber",
    "filmy_dt"
  )
)
```

```{r load-functions}
# Define functions specific to this file
plot_chamber_rh <- function (data, time_limits) {
  ggplot(data, aes(x = date_time, y = 0.01*rh)) +
    geom_line() +
    scale_x_datetime(limits = time_limits) +
    scale_y_continuous(label = label_percent(accuracy = 1)) +
    labs(y = "Relative humidity") +
    theme(
      legend.position = "none",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank())
}

@@

```{r gametophyte-dt-2013, fig.cap = ""}
# Calculate mean start/end times for gametophyte DT groups
gameto_mean_times <-
  filmy_dt %>%
  filter(generation == "gametophyte") %>%
  select(individual, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", -individual) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(individual) %>%
  filter(n > 1) %>%
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time") %>%
  # Cluster starts times into groups (here, I know there are 10 groups)
  add_clusters("pre", 10) %>%
  mutate(cluster = as.factor(cluster)) %>%
  group_by(cluster) %>%
  summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop")

@@

ggsave(here::here("results/S1_fig.pdf"))
```

**Figure S1**. Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes. MgNO3 was used to maintain constant humidity at \circa{ }58%. All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days by accident.

@@

```{r sporophyte-dt-2013-LiCl, fig.cap = ""}
# Get mean times of each cluster of DT test samples for the selected drying salt
sporo_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
sporo_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)

@@

ggsave(here::here("results/S2_fig.pdf"))
```

@@

>>>>

---
title: "SI"
bibliography: 
  - references.bib
  - references_other.yaml
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research-si.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

@@

```{r load-data, cache = FALSE}
tar_load(
  names = c(
    "filmy_dt_chamber",
    "filmy_dt",
    "filmy_habit",
    "rel_water_indiv",
    "rel_water_species_means"
  )
)
```

```{r load-functions, cache = FALSE}
# Define functions specific to this file
plot_chamber_rh <- function (data, time_limits) {
  ggplot(data, aes(x = date_time, y = 0.01*rh)) +
    geom_line() +
    scale_x_datetime(limits = time_limits) +
    scale_y_continuous(label = label_percent(accuracy = 1)) +
    labs(y = "Relative humidity") +
    theme(
      legend.position = "none",
      axis.title.x = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank())
}

@@

```{r gametophyte-dt-2013, fig.height = 4.5, fig.width = 7, fig.cap = ""}
# Calculate mean start/end times for gametophyte DT groups
gameto_mean_times <-
  filmy_dt %>%
  filter(generation == "gametophyte") %>%
  select(individual, contains("time_")) %>%
  pivot_longer(names_to = "condition", values_to = "date_time", -individual) %>%
  filter(!is.na(date_time)) %>%
  mutate(condition = str_remove_all(condition, "time_") %>%
           str_replace_all("30min", "post")) %>%
  filter(condition %in% c("pre", "post")) %>%
  add_count(individual) %>%
  filter(n > 1) %>%
  select(-n) %>%
  pivot_wider(names_from = "condition", values_from = "date_time") %>%
  # Cluster starts times into groups (here, I know there are 10 groups)
  add_clusters("pre", 10) %>%
  mutate(cluster = as.factor(cluster)) %>%
  group_by(cluster) %>%
  summarize(pre = mean(pre), post = mean(post), n = n(), .groups = "drop")

@@

# ggsave(here::here("results/S1_fig.png"))
```

**Figure S1**. Conditions inside desiccation chamber and duration of desiccation treatment for gametophytes. MgNO~3~ was used to maintain constant humidity at \circa{ }58%. All gametophytes were treated to desiccation for 2 days, except for group 8, which was treated for four days unintentionally.

\clearpage

@@

```{r sporophyte-dt-2013-LiCl, fig.height = 4.5, fig.width = 7, fig.cap = ""}
# Get mean times of each cluster of DT test samples for the selected drying salt
sporo_clusters_2013 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2013")), "LiCl", 3)
sporo_clusters_2014 <- cluster_sporo_times(filter(sporo_times, str_detect(dataset, "2014")), "LiCl", 1)

@@

# ggsave(here::here("results/S2_fig.png"))
```

@@

\clearpage

```{r water-content-prep}
# Define color-blind palette
cbPalette <- c(grey = "#999999", gold = "#E69F00", skyblue = "#56B4E9", 
               green = "#009E73", yellow = "#F0E442", blue = "#0072B2", 
               red = "#D55E00", purple = "#CC79A7")

# Set colors, shapes for desiccation intensity
salt_cols <- c(
  "NaCl" = cbPalette[["blue"]], 
  "MgNO3" = cbPalette[["yellow"]], 
  "LiCl" = cbPalette[["red"]], 
  "Control" = "white")

salt_shapes <- c(
  "NaCl" = 21, # circle
  "MgNO3" = 24, # triangle
  "LiCl" = 23, # diamond
  "Control" = 22 # square
  )
```

```{r water-content-full, fig.height = 4.5, fig.width = 7, fig.cap = ""}
# Plot only species with weights recorded during recovery
species_with_rec <-
  rel_water_species_means %>%
  filter(rec_time %in% c("30min", "24hr", "48hr")) %>%
  pull(species) %>%
  unique()

rel_water_species_means_with_rec <-
  rel_water_species_means %>%
  filter(species %in% species_with_rec)

pd <- position_dodge(.5)

# Define color-blind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Set colors for desiccation intensity
salt_cols <- c("NaCl" = cbPalette[7], "MgNO3" = cbPalette[5], "LiCl" = cbPalette[6], "Control" = "white")

ggplot(rel_water_species_means_with_rec, aes(x = rec_time, y = rel_water, group = interaction(dry_time, salt))) +
  geom_errorbar(
    aes(ymin = rel_water - sd, ymax = rel_water + sd),
    color = "grey50",
    width = 0,
    position = pd) +
  geom_line(position = pd, aes(linetype = dry_time), color = "grey30") +
  geom_point(aes(shape = salt, fill = salt), position = pd) +
  scale_linetype_manual(name = "Desiccation Time",
                        values = c("solid", "dashed"),
                        breaks = c("15", "2"),
                        labels = c("15 days", "2 days")) +
  scale_shape_manual(name = "Desiccation Intensity",
                     values = salt_shapes,
                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
  scale_fill_manual(name = "Desiccation Intensity",
                    values = salt_cols, 
                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
  scale_y_continuous(label = scales::percent) +
  facet_wrap(vars(species)) +
  labs(x = "Condition",
       y = "Relative water content") +
  theme_bw() +
  theme(
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    legend.box="vertical", 
    legend.margin=margin()
  )

# ggsave(here::here("results/S_fig_water_content_full.png"))
```

**Figure SXX**. Relative water content (RWC) of sporophytes during desiccation tolerance test. RWC was not measured for all combinations of species and desiccation treatments; only species with complete data shown.

\clearpage

```{r water-content-desiccated, fig.height = 4.5, fig.width = 7, fig.cap = ""}
# Plot relative water content in desiccated state for all species
rel_water_species_means_desic <<-
rel_water_species_means %>%
  filter(rec_time == "desiccated") %>%
  select(-rec_time)

# Set dodge spacing, colors
pd <- position_dodge(.7)

ggplot(rel_water_species_means_desic, aes(x = rel_water, y = species, group = salt)) +
  geom_errorbar(
    aes(xmin = rel_water - sd, xmax = rel_water + sd),
    color = "grey80",
    width = 0.2,
    position=position_dodge(width=0.5)) +
  geom_point(aes(fill = salt, shape = salt), size = 1.25, position=position_dodge(width=0.5)) +
  scale_x_continuous(label = scales::percent) +
  scale_shape_manual(name = "Desiccation Intensity",
                     values = salt_shapes,
                     breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
                     labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) + 
  scale_fill_manual(name = "Desiccation Intensity",
                    values = salt_cols, 
                    breaks = c("NaCl", "MgNO3",  "LiCl", "Control"),
                    labels = c("-38 MPa", "-86 MPa", "-282 Mpa", "Control")) +
  labs(
    x = "Relative Water Content",
    y = "Species",
    color = "Desiccation\nintensity") +
  theme_bw() +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position = "bottom"
  ) +
  facet_grid(vars(dry_time), labeller = as_labeller(c(`2` = "2 days", `15` = "15 days")))

# ggsave(here::here("results/S_fig_water_content_desic.pdf"))
```

**Figure SXX**. Relative water content (RWC) of sporophytes after desiccation by duration of desiccation treatment (2 or 15 days). RWC was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not RWC of zero).

\clearpage

```{r sporo-dt-controls, fig.height = 8, fig.width = 7, fig.cap = ""}
# Plot sporophyte DT controls

# Prepare data for plotting
sporo_control_plot_data <-
  # Start with DT data
  filmy_dt %>%
  filter(salt == "Control", generation == "sporophyte") %>%
  # Filter out individuals with low pre-treatment yields
  filter(!is.na(yield_pre), yield_pre > 400) %>%
  # Calculate recovery by individual
  verify(yield_pre > 0) %>%
  assert(not_na, yield_pre) %>%
  select(species, salt, dry_time, individual, dataset, contains("yield")) %>%
  select(-yield_72hr, -contains("yield_dry")) %>%
  pivot_longer(names_to = "rec_time", values_to = "yield_recover", matches("30|24|48")) %>%
  mutate(
    rec_time = str_remove_all(rec_time, "yield_"),
    recovery = yield_recover / yield_pre) %>%
  select(-contains("yield")) %>%
  # Calculate mean recovery by species and dataset
  mutate(rec_time = factor(rec_time, levels = c("30min", "24hr", "48hr"))) %>%
  group_by(species, salt, dry_time, rec_time, dataset) %>%
  summarize(
    recovery_mean = mean(recovery, na.rm = TRUE),
    sd = sd(recovery, na.rm = TRUE),
    n = n(),
    .groups = "drop") %>%
  rename(recovery = recovery_mean) %>%
  # Add growth habit data
  left_join(filmy_habit, by = "species") %>%
  assert(not_na, habit) %>%
  mutate(
    habit = str_replace_all(habit, "_", " ") %>% 
      str_to_sentence() %>%
      factor(levels = c("Terrestrial", "Saxicolous", "Low epiphyte", "High epiphyte"))) %>%
  # Add dataset year
  mutate(year = str_split(dataset, "_") %>% map_chr(1)) %>%
  # Abbreviate species names to "G. species"
  mutate(
    species = str_replace_all(species, "[a-z]+_", ". "),
    species = glue("*{species}*")) %>%
  # Add year to species names with multiple datasets
  add_count(species, name = "n_species") %>%
  mutate(species = ifelse(n_species > 6, paste(species, year), species)) %>%
  select(-n_species) %>%
  # Only keep rows with recovery
  filter(!is.na(recovery))

# Make plot
ggplot(sporo_control_plot_data, aes(x = rec_time, y = recovery, shape = dry_time, fill = salt)) +
  geom_point(position = pd, size = 1.5, fill = "white") +
  geom_errorbar(
    aes(ymin = recovery - sd, ymax = recovery + sd), 
    colour = "grey50", width = 0, 
    position = pd) +
  facet_wrap(vars(habit, species)) +
  scale_shape_manual(name = "Desiccation Time",
                        values = c(24, 21),
                        breaks = c("15", "2"),
                        labels = c("15 days", "2 days")) +
  scale_y_continuous(label = label_percent(accuracy = 1)) +
  scale_x_discrete(limits = c("30min","24hr","48hr"),
                   labels = c("0.5", "24", "48")) +
  xlab("Recovery Time (hr)") +
  ylab("Recovery (%)") +
  theme_bw() +
  legend_theme +
  theme(legend.position ="bottom",
        panel.grid.minor = element_blank(), 
        panel.grid.major.x = element_blank(),
        strip.text = element_markdown()
)
```

**Figure SXX**. Recovery of controls (samples kept on moist tissues) for desiccation tolerance test of filmy fern sporophytes from Moorea, French Polynesia. Recovery was measured by comparing maximum photochemical yield of photosystem II before desiccation treatment with values at 30 min, 24 hr, and 48 hr following desiccation treatment (see Methods). Recovery was not measured for all combinations of species and desiccation treatments; missing points indicate data not collected (not recovery of 0%).
